Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio

### 2017-05-15 paper.tex
Added sth. to the intro and fixed a typo.
some changes
Saved Thiemos changes.

\author{\IEEEauthorblockN{Maximilian Stiefel\IEEEauthorrefmark{1}, Elmar van Rijnswou\IEEEauthorrefmark{1}\\ 
Carlos Pérez-Penichet\IEEEauthorrefmark{1} Ambuj Varshney\IEEEauthorrefmark{1}, 
Christian Rohner\IEEEauthorrefmark{1} and Thiemo Voigt\IEEEauthorrefmark{2}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Uppsala University}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala Univerisyt and RISE SICS}}

@@

The received signal can be interpreted as
\begin{equation}
	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

@@

We design our backscatter transmitter to be used with ambient
television signals.  We design the transmitter for ambient
television signal present in the band with  
center frequency of \SI{626}{\mega\hertz}. Crucial to the design
of the transmitter is the implementation of an antenna, as
the antennas are known to be  frequency selective. To this end, we design an
antenna on an unetched board made of FR4 substrate commonly
used to design PCBs. The board acts like the ground plane, at the center
of the board we have a wire acting as a monopole antenna element. 
We have verified using a vector network analyzer (VNA), that the antenna
can operate from frequency range most commonly
used for television broadcast, 
from \SI{245}{\mega\hertz} to \SI{626}{\mega\hertz}.

>>>>

\author{\IEEEauthorblockN{Maximilian Stiefel\IEEEauthorrefmark{1}, Elmar van Rijnswou\IEEEauthorrefmark{1}\\ 
Carlos Pérez-Penichet\IEEEauthorrefmark{1} Ambuj Varshney\IEEEauthorrefmark{1}, 
Christian Rohner\IEEEauthorrefmark{1} and Thiemo Voigt\IEEEauthorrefmark{2}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Uppsala University}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala University and RISE SICS}}

@@

Quadrature demodulation is the major feature of the \textit{RTL2832U},  which makes it possible to decode a digitally amplitude- and phase-shifted signal. The received signal (behind the tuner at the input of the \textit{RTL2832U}) can be interpreted as
\begin{equation}
	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

@@

We design our backscatter transmitter to be used with ambient
television signals.  We design the transmitter for ambient
television signal present in the band with  
center frequency of \SI{626}{\mega\hertz}. Crucial to the design
of the transmitter is the implementation of an antenna, as
the antennas are known to be  frequency selective. To this end, we design an
antenna on an unetched board made of FR4 substrate commonly
used to design PCBs. The board acts like the ground plane, at the center
of the board we have a wire acting as a monopole antenna element. 
The antenna has a quite good reflection coefficient \ensuremath{S_{11}} 
of -17 dB at 626 MHz, which is the frequency aimed for. This means, that at 
this frequency the antenna absorber more energy form the electromagnetic field compared to another band. A vector network analyzer (VNA) has been used to analyze antenna performance.
